{\normalsize Seminar: Formal Specification} \\[1ex] Security of Multithreaded Programs by Compilation\cite{Barthe07}

### 2012-01-26 paper/paper.tex
beautified some more

At this point a first simplification can be made. Looking a state $s \in ConcState$ we can
first extract the active threads (\texttt{s.act}) by taking the domain of the first
component. According to a security environment we can classify these threads,
with respect to their current program point (\texttt{s.pc(tid)} where $tid \in Thread$),
into \textit{low} threads iff the current program point is low. In \textit{high}
threads iff the current program point is high. In \textit{always high} threads
iff the current program point is always high and in \textit{hidden} threads, iff
the current program point is high but not always high.

@@

To complete our multithreaded setup we need a scheduler. The scheduler will
operate on histories. A history is a list of pairs $(tid, l$), where $tid \in
Thread$ and $l \in Level$. In this history all threads chosen by the scheduler
are recorded.

At this point no concrete scheduler is defined to make the framework applicable
for a wide class of schedulers. A scheduler is in this class iff it can be
modeled as a function $pickt: ConcState \times History \rightharpoonup Thread$
which satisfies the following constraints:

@@

To define what noninterference means we need a notion of excution. It is assumed
that we have a sequential execution relation $\leadsto_{seq} \subseteq SeqState \times
SeqState$. The one step execution for the multithreaded language $\leadsto_{seq}
\subseteq (ConcState \times History) \times (ConcState \times History)$ is defined
by the following rules.

\begin{figure}
\begin{align*}
\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] \in SeqIns \\
\langle s(ctid), s.gmem\rangle \leadsto_{seq} \sigma,\mu \\ \sigma.pc \neq \mathtt{exit}}
{s,h \leadsto_{conc} s[lst(ctid) := \sigma, gmem := \mu], \langle ctid, se(i)\rangle :: h}
\end{align*}

\begin{align*}
\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] \in SeqIns \\
\langle s(ctid), s.gmem\rangle \leadsto_{seq} \sigma,\mu \\ \sigma.pc = \mathtt{exit}}
{s,h \leadsto_{conc} s[lst := lst \\ ctid, gmem := \mu], \langle ctid, se(i)\rangle :: h}
\end{align*}

@@

The third rule introduces the dynamic creation of new threads. In this rule the function $fresh_{l}$
takes a set of thread identifiers and returns a new thread identifier at level $l$\footnote{It is
assumed that $fresh_l \neq fresh_{l'} Leftrightarrow l \neq l'$} and $\lambda_{init} : \mathcal{P}
\rightarrow LocState$ produces an initial state with the program pointer at the given program point.
If the current instruction at program point $i$ is the \texttt{start} instruction, a new thread
identifier is picked w.r.t. the security level of $i$. The program counter of the current thread
is increased by one, to step to the next instruction. The resulting concurrent state includes the
updated state of the current thread and the initial state for the new thread. The global memory
is not modified and the history is extended with the current thread id and the security level of $i$.

Based on this an evaluation relation $\Downarrow_{conc} \subseteq (ConcState \times History) \times
GMemory)$ is defined by

@@

\todo{beautify this}
\begin{align*}
instr &::= binop\ op && \text{binary operation with values from stack} \\
&|\ push\ n &&  \text{push value on the stack} \\
&|\ load\ x && \text{push value of variable on the stack} \\
&|\ store\ x && \text{store first element of the stack in x} \\
&|\ goto\ j\ |\ ifeq\ j && \text{un-/conditional jump to j} \\
&|\ start\ j && \text{create a new thread starting in j} \\
\end{align*}

@@

Besides this concrete language we need to define a type system to enforce noninterference
according to \ref{sec:typesystem}. The local types are defined by a stack of security levels
$LType = Stack(Level)$ and $t_{init}$ with the empty stack. The typing rules defined in
Figure~\ref{fig:multithreaded-typing-rules} need to be extended with rules for the concrete instructions
of the assembly language. In this summary I only explain two of them, the rest follows
in a similar manner.

\todo{beautify me}
\begin{figure}
\begin{align*}
\inferrule{P[i] = store\ x \\ se(i) \sqcup k \leq \Gamma(x)}
{se, i \vdash_{seq} k :: st \Rightarrow st}
\end{align*}

\begin{align*}
\inferrule{P[i] = ifeq\ j \\ \forall j' \in reg(i), k \leq se(j')}
{se, i \vdash_{seq} k :: st \Rightarrow lift_k(st)}
\end{align*}
\caption{Excerpt of typing rules for the assembly language}
\label{fig:transfer-rules}
\end{figure}

@@

\begin{align*}
e ::= n\ |\ x\ |\ e\ op\ e && c\ ::=\ x\ :=\ e\ |\ c;c\ |\ if\ e\ then\ c\ else\ c\ |\ while\ e\ do\ c\ |\ fork(c)
\end{align*}

@@

\begin{align*}
c\ ::=\ [x\ :=\ e]^n\ |\ c;c\ |\ [if\ e\ then\ c\ else\ c]^n\ |\ [while\ e\ do\ c]^n\ |\ [fork(c)]^n
\end{align*}

@@

\todo{use an other package to display these rules}
\begin{figure}
\begin{align*}
\inferrule[L-COND]{\vdash e : L \\ \vdash_\alpha c : E \\ \vdash_\alpha c' : E}
{\vdash_\alpha [if \ e \ then \ c \ else \ c']^n_\alpha : E}
\end{align*}

\begin{align*}
\inferrule[H-COND]{\vdash e : H \\ \vdash_\bullet c : E \\ \vdash_\bullet c' : E}
{\vdash_\bullet [if \ e \ then \ c \ else \ c']^n_\bullet : E}
\end{align*}

\begin{align*}
\inferrule[TOP-H-COND]{\vdash e : H \\ \vdash_\bullet c : E \\
\vdash_\bullet c' : E \\ E = lift_H(E, sregion(n))}
{\vdash_\circ [if \ e \ then \ c \ else \ c']^n_\bullet : E}
\end{align*}
\caption{Typing rules for \texttt{if} on source level.}
\label{fig:typing-rules-if}
\end{figure}

The first rule L-COND from Figure \ref{fig:typing-rules-if} says that if we branch on a low
guard, everything depends on the security level of $c$ and $c'$. The rule H-COND covers the
case where we have a high guard, in this case the control dependence region has to be marked
high. The rule TOP-H-COND is the interesting one. Because of the preceding rules we can not
be part of branch with a low guard, therefore we are in the outermost high branch. The premise
$E' = lift_H(E, sregion(n))$ reads as: For all labels in the control dependence region $E$ is
defined as $E'(n) = H \sqcup E(n)$ and $E'(n) = E(n)$ for all other labels.

>>>>

At this point a first simplification can be made. Looking a state $s \in \text{ConcState}$ we can
first extract the active threads (\texttt{s.act}) by taking the domain of the first
component. According to a security environment we can classify these threads,
with respect to their current program point (\texttt{s.pc(tid)} where $tid \in \text{Thread}$),
into \textit{low} threads iff the current program point is low. In \textit{high}
threads iff the current program point is high. In \textit{always high} threads
iff the current program point is always high and in \textit{hidden} threads, iff
the current program point is high but not always high.

@@

To complete our multithreaded setup we need a scheduler. The scheduler will
operate on histories. A history is a list of pairs $(tid, l$), where $tid \in
\text{Thread}$ and $l \in \text{Level}$. In this history all threads chosen by the scheduler
are recorded.

At this point no concrete scheduler is defined to make the framework applicable
for a wide class of schedulers. A scheduler is in this class iff it can be
modeled as a function $pickt: \text{ConcState} \times \text{History} \rightharpoonup \text{Thread}$
which satisfies the following constraints:

@@

To define what noninterference means we need a notion of excution. It is assumed
that we have a sequential execution relation $\leadsto_{seq} \subseteq \text{SeqState} \times
\text{SeqState}$. The one step execution for the multithreaded language $\leadsto_{seq}
\subseteq (\text{ConcState} \times \text{History}) \times (\text{ConcState} \times \text{History})$ is defined
by the following rules.

\begin{figure}
\begin{align*}
\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] \in \text{SeqIns} \\
\langle s(ctid), s.gmem\rangle \leadsto_{seq} \sigma,\mu \\ \sigma.pc \neq \mathtt{exit}}
{s,h \leadsto_{conc} s[lst(ctid) := \sigma, gmem := \mu], \langle ctid, se(i)\rangle :: h}
\end{align*}

\begin{align*}
\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] \in \text{SeqIns} \\
\langle s(ctid), s.gmem\rangle \leadsto_{seq} \sigma,\mu \\ \sigma.pc = \mathtt{exit}}
{s,h \leadsto_{conc} s[lst := lst \\ ctid, gmem := \mu], \langle ctid, se(i)\rangle :: h}
\end{align*}

@@

The third rule introduces the dynamic creation of new threads. In this rule the function $fresh_{l}$
\footnote{It is assumed that $fresh_l \neq fresh_{l'} Leftrightarrow l \neq l'$}
takes a set of thread identifiers and returns a new thread identifier at level $l$ and $\lambda_{init} : \mathcal{P}
\rightarrow \text{LocState}$ produces an initial state with the program pointer at the given program point.
If the current instruction at program point $i$ is the \texttt{start} instruction, a new thread
identifier is picked w.r.t. the security level of $i$. The program counter of the current thread
is increased by one, to step to the next instruction. The resulting concurrent state includes the
updated state of the current thread and the initial state for the new thread. The global memory
is not modified and the history is extended with the current thread id and the security level of $i$.

Based on this an evaluation relation $\Downarrow_{conc} \subseteq (\text{ConcState} \times \text{History}) \times
\text{GMemory})$ is defined by

@@

\begin{align*}
instr &::= \text{binop}\ op && \text{binary operation with values from stack} \\
&|\ \text{push}\ n &&  \text{push value on the stack} \\
&|\ \text{load}\ x && \text{push value of variable on the stack} \\
&|\ \text{store}\ x && \text{store first element of the stack in $x$} \\
&|\ \text{goto}\ j\ |\ \text{ifeq}\ j && \text{un-/conditional jump to $j$} \\
&|\ \text{start}\ j && \text{create a new thread starting in $j$} \\
\end{align*}

@@

Besides this concrete language we need to define a type system to enforce noninterference
according to section \ref{sec:typesystem}. The local types are defined by a stack of security levels
$\text{LType} = \text{Stack}(\text{Level})$ and $\textsc{t}_{init}$ with the empty stack. The typing rules defined in
figure~\ref{fig:multithreaded-typing-rules} need to be extended with rules for the concrete instructions
of the assembly language. In this summary I only explain two of them, the rest follows
in a similar manner.

\begin{figure}
\begin{minipage}{.5\textwidth}
\begin{align*}
\inferrule{P[i] = \text{store}\ x \\ se(i) \sqcup k \leq \Gamma(x)}
{se, i \vdash_{seq} k :: st \Rightarrow st}
\end{align*}
\end{minipage}
\begin{minipage}{.5\textwidth}
\begin{align*}
\inferrule{P[i] = \text{ifeq}\ j \\ \forall j' \in reg(i), k \leq se(j')}
{se, i \vdash_{seq} k :: st \Rightarrow lift_k(st)}
\end{align*}
\end{minipage}
\caption{Excerpt of typing rules for the assembly language.}
\label{fig:transfer-rules}
\end{figure}

@@

\begin{align*}
e ::= n\ |\ x\ |\ e\ op\ e && c ::= x\ :=\ e\ |\ c;c\ |\ \text{if}\ e\ \text{then}\ c\ \text{else}\ c\ |\ \text{while}\ e\ \text{do}\ c\ |\ \text{fork}(c)
\end{align*}

@@

\begin{align*}
c ::= [x\ :=\ e]^n\ |\ c;c\ |\ [\text{if}\ e\ \text{then}\ c\ \text{else}\ c]^n\ |\ [\text{while}\ e\ \text{do}\ c]^n\ |\ [\text{fork}(c)]^n
\end{align*}

@@

\todo{use an other package to display these rules}
\begin{figure}
\begin{align*}
\inferrule[L-Cond]{\vdash e : L \\ \vdash_\alpha c : E \\ \vdash_\alpha c' : E}
{\vdash_\alpha [\text{if} \ e \ \text{then} \ c \ \text{else} \ c']^n_\alpha : E}
\end{align*}

\begin{align*}
\inferrule[H-Cond]{\vdash e : H \\ \vdash_\bullet c : E \\ \vdash_\bullet c' : E}
{\vdash_\bullet [\text{if} \ e \ \text{then} \ c \ \text{else} \ c']^n_\bullet : E}
\end{align*}

\begin{align*}
\inferrule[Top-H-Cond]{\vdash e : H \\ \vdash_\bullet c : E \\
\vdash_\bullet c' : E \\ E = lift_H(E, sregion(n))}
{\vdash_\circ [\text{if} \ e \ \text{then} \ c \ \text{else} \ c']^n_\bullet : E}
\end{align*}
\caption{Typing rules for \texttt{if} on source level.}
\label{fig:typing-rules-if}
\end{figure}

The first rule \textsc{L-Cond} from Figure \ref{fig:typing-rules-if} says that if we branch on a low
guard, everything depends on the security level of $c$ and $c'$. The rule \textsc{H-Cond} covers the
case where we have a high guard, in this case the control dependence region has to be marked
high. The rule \textsc{Top-H-Cond} is the interesting one. Because of the preceding rules we can not
be part of branch with a low guard, therefore we are in the outermost high branch. The premise
$E' = lift_H(E, sregion(n))$ reads as: For all labels in the control dependence region $E$ is
defined as $E'(n) = H \sqcup E(n)$ and $E'(n) = E(n)$ for all other labels.
